### 2013-07-06 appendix.tex
Initial commit

% This is appendix.tex
% This file is for appendices.

%%% If you want to include appendices, delete
%%% the "\endinput" command directly below!

\endinput % Delete this command if you want to include appendices.

\appendix

% Add additional appendices by replicating the line below.

\section{Title of Appendix}

Appendices may appear after the paper proper.  Appendices may hold
extra information that would interrupt the flow of the paper and that
is not absolutely necessary for the reader to appreciate the work.
For example, a large number of related figures or a mathematical
derivation could go nicely in an appendix.

### 2013-07-06 bibliography.tex
Initial commit

% This is bibliography.tex

% If you don't have a bibliography, use the "\endinput" command here.
% \endinput

\begin{thebibliography}{99}

% Write your bibliography here using the \bibitem{} command. 
% Be sure to strictly follow your bibliography formatting system (eg. Vancouver, Harvard, Oxford, or IEEE-system).

%Example in shown below:

\bibitem{digital imaged based tree measurement for forest inventory}
Juujärvi J, Heikkonen J, Brandt S, Lampinen J. Digital Image Based Tree Measurement for Forest Inventory. Proceedings of SPIE. 1998; Volume 3522: 114-123.

\end{thebibliography}

### 2013-07-06 paper.tex
Initial commit

%This is paper.tex

%Write the contents of your paper here.

\newcommand{\degree}{$^\circ$}

\section{Introduction}
Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory. The use of this information is important to companies who need to make smart, profitable decisions but also need to be certain that the choices they make now are maintable in the future. However this information also benefits climate modellers, environmental groups and governments who want to be able to sustain ecosystems and plan restorations. To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis. Most methods used today to measure individual trees are often done manually with a variety of tools. The use of these analogue approaches is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}. SilviaTerra, however, is a company willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results. This is achieved through computer exactitude combined with statistical analysis. Recently a new project was started, with this research paper developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.

The measurements required for forest inventory can be achieved through a number of methods. One of them is using a so called biltmore stick. A biltmore stick is similar to a normal ruler but has been specially modified for measuring trees. When standing at a certain distance the user can discern the height of the tree by holding the biltmore stick up to their eye and reading the corresponding number. The biltmore stick can also easily be used to measure the DBH of the tree. Other ways of measuring tree volume include using a combination of lasers and inclinometers or by having a person manually climb up the tree and drop a tape measure down to measure the height. Despite these relatively simpe ways of measuring they often leave room for error. That’s why a more common way of getting a precise reading is to cut down trees in the area and then physically measure them. After the relevant data is collected one applies a stem-profile taper equation that is specific to the species of tree measured. These taper functions are based on numerous measurements and have been refined over the years. The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions. This again is cost prohibitive and only so much of a forest can be accounted for in the taper function. Even when one manages to create a new taper function, it is still far from perfect as all trees, even from the same species, differ in shape.

This project aims to solve the described problem by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation. This is supposed to then give a accurate result, cheaply and non-destructively. All within the reaches of a mobile phone.

\section{Background}

\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.65\textwidth]{main.png}
	  	\caption{Homescreen}
  		\label{main}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.65\textwidth]{input.png}
	  	\caption{User input}
  		\label{input}
	\endminipage\hfill
\end{figure}
The first thing a person does when accessing the mobile application is to log into the system with their user name and password. After that they are prompted with a simple menu where they can either choose to add new trees to their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}. The data can then later be used with SilviaTerra's other products. If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, \ref{input}. 
\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.75\textwidth]{dbh.png}
	  	\caption{DBH picture}
	  	\label{dbh}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.75\textwidth]{bole.png}
	  	\caption{Bole picture}
	  	\label{bole}
	\endminipage\hfill
\end{figure}

After that they take two pictures. The interface is quite simple with a cross-hair to show where to point the camera. The first picture is of the actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{bole}. Between these two pictures the camera will have to be tilted vertically to capture the bole. The angle between these two camera positions is recorded through the internal gyroscope. After this stage the user's role in the application is done and the pictures is sent to one of SilviaTerra's web servers. Here horizontal lines that run parallel with the bottom of the image are drawn across the entire picture. These are drawn at different pixel heights with equal pixel distance in between. The reason this is done is because computer vision is still an experimental technology and is not fully capable yet of being to detect the edges of trees with a background of forestry. Seeing as this is a commercial product, it would be quite disadvantageous if errors were commonplace. So instead the images with the newly drawn lines are then forwarded to Amazon's Mechanical Turk service. Just like the name suggests Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of a automated chess board. The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed. Usually the tasks are ones that are difficult for computers but easy for humans. SilviaTerra creates tasks where people manually click where the previously drawn lines, cross the edges of the tree. Ones of the points of intersection are identified, their corresponding pixel coordinates are the sent back to the web server. The pixel coordinates along with information from the phone is piped into a program where the volume is then calculated. It is this program that is the main topic of this research paper.

\section{Method}
Once all the required information has been retrieved it is piped into the program that will be calculating the volume. The dataset that is used comprises information about the phone and the results from Mechanical Turk. It is as follows:
\begin{itemize}
	\item DBH length
	\item Horizontal and vertical angle of view
	\item Horizontal and vertical image resolution
	\item Pitch of phone when images are captured
	\item Left and right pixel coordinates from Mechanical Turk
\end{itemize}

\subsection{Distance to tree}
The DBH is used as a reference to find all the other measurements. The first unknown that must be discovered is the distance to the tree. This distance is parallel with the ground. The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels betwee them.
\begin{figure}[htp]
\centering
{
	\includegraphics[scale=0.3]{horizontal_triangle.pdf}
	\caption{Birds eye view of horizontal plane}
	\label{horizontal_triangle}
}
\end{figure}
 As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constructed with the horizontal angle of view as the vertex angle. The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree. To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image. The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, contriving the angle $\theta$ which is exactly half of the angle of view and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore producing a right angle triangle that can be solved. Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.

\subsection{Heights and widths} 
\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.9\textwidth]{triangle1.pdf}
	  	\caption{DBH picture}
	  	\label{triangle1}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.9\textwidth]{triangle2.pdf}
	  	\caption{Bole picture}
	  	\label{triangle2}
	\endminipage\hfill
\end{figure}

The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale which is the ratio between degrees and pixels in the vertical axis is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane. With this method it is possible to find out the angle to each pixel coordinate, $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the second image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them. This is accomplished by employing the method for calculating the distance to the tree, in reverse. In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement. To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume. The current implemented method of doing this is by integrating all the points using the trapezoidal rule. All the widths are divided by 2 to get the radius. Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height. The basal area is plotted along with their corresponding heights. The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$. The area between all the points is the summed together.

This whole process is looped through for every identified pixel coordinate set until the width and height is known for all of them. Once that is done, a new loop is created for calculating the total volume.

\subsection{Simulation}
To test this method a \emph{virtual tree} was constructed. This tree was made out of blue tape and tapered off so that the program could be thoroughly tested. Along the tape points were marked out and the height to them plus the width between corresponding points were measured. Just like in the real application, two pictures were taken. One of the DBH and one of the top of the tree. Each set of points was then identified on the pictures and measured in pixels. The parameters were then passed into the program and the results were cross-referenced with the real life measurements.

\newpage

\section{Results}
\begin{table}[h!]
	\begin{center}
		\begin{tabular}{| l c c c c c c r |}		
		\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		\hline
		0 		& 0 	& 0 	& 0 	& 0 	& 0 	& 0 	& 0 	\\
		103 	& 96 	& 100 	& 103 	& 106 	& 113 	& 116 	& 123 	\\
		115 	& 109 	& 113 	& 116 	& 118 	& 125 	& 128 	& 135 	\\
		125 	& 125 	& 129	& 132 	& 134 	& 140 	& 144 	& 150 	\\
		159 	& 142 	& 147 	& 150 	& 151 	& 158 	& 161 	& 167 	\\
		166 	& 158 	& 163 	& 166 	& 167 	& 174 	& 175 	& 183 	\\
		183 	& 173 	& 179 	& 181	& 182 	& 189 	& 192 	& 198 	\\
		\hline
		\end{tabular}
		\caption{Heights over DBH}
		\label{heights}
	\end{center}
\end{table}

\begin{table}[h!]
	\begin{center}
		\begin{tabular}{| l c c c c c c r |}
		\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		\hline
		12       & 11.7      & 12.1      & 12.2      & 12.1      & 12.3      & 12.6      & 11.9      \\
		27       & 26.8      & 28.0      & 28.4      & 28.2      & 28.5      & 28.4      & 28.1      \\
		37       & 36.4      & 37.5      & 38.0      & 37.7      & 38.2      & 38.1      & 37.7      \\
		44       & 42.4      & 43.6      & 44.4      & 44.0      & 44.4      & 44.0      & 44.4      \\
		52       & 51.3      & 52.7      & 53.5      & 53.0      & 53.4      & 52.9      & 53.2      \\
		61       & 60.9      & 62.2      & 63.1      & 62.1      & 62.8      & 62.5      & 61.1      \\
		59       & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      \\
		\hline
		\end{tabular}
		\caption{Widths over DBH}
		\label{widths}
    \end{center}
\end{table}

\begin{table}[h!]
	\begin{center}
    	\begin{tabular}{| l c c c c c c r |}
    	\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401    & 445624    \\
		\hline
		\end{tabular}
		\caption{Total volume over DBH}
		\label{volumes}
    \end{center}
\end{table}

### 2013-07-06 paper.tex
breaking lines to 80 chars

@@
 \section{Introduction}
-Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory.
-The use of this information is important to companies who need to make smart, profitable decisions but also need to be certain that the choices they make now are maintable in the future.
-However this information also benefits climate modellers, environmental groups and governments who want to be able to sustain ecosystems and plan restorations.
-To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.
-Most methods used today to measure individual trees are often done manually with a variety of tools.
-The use of these analogue approaches is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}.
-SilviaTerra, however, is a company willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results.
-This is achieved through computer exactitude combined with statistical analysis.
-Recently a new project was started, with this research paper developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.
+Tree volume measurement is one of the many parameters in documenting and 
+profiling trees and is at the center of forest inventory. The use of this 
+information is important to companies who need to make smart, profitable 
+decisions but also need to be certain that the choices they make now are 
+maintable in the future. However this information also benefits climate 
+modellers, environmental groups and governments who want to be able to sustain 
+ecosystems and plan restorations. To be able to accomplish this in forestry, 
+it is a necessity to be able to collect large amounts of data easily and on a 
+regular basis. Most methods used today to measure individual trees are often 
+done manually with a variety of tools. The use of these analogue approaches 
+is expensive, time consuming and in most cases the data collected doesn't meet 
+today's standards \cite{digital imaged based tree measurement for forest inventory}. 
+SilviaTerra, however, is a company willing to change that and have already 
+developed several tools for scanning, analyzing and mapping areas of forestry 
+to give accurate results. This is achieved through computer exactitude combined 
+with statistical analysis. Recently a new project was started, with this 
+research paper developing the mathematical back end, to allow users to easily 
+get a correct volume measurement by only using their smart phone.
 
-The measurements required for forest inventory can be achieved through a number of methods.
-One of them is using a so called biltmore stick.
-A biltmore stick is similar to a normal ruler but has been specially modified for measuring trees.
-When standing at a certain distance the user can discern the height of the tree by holding the biltmore stick up to their eye and reading the corresponding number.
+The measurements required for forest inventory can be achieved through a number 
+of methods. One of them is using a so called biltmore stick. A biltmore stick 
+is similar to a normal ruler but has been specially modified for measuring trees. 
+When standing at a certain distance the user can discern the height of the tree 
+by holding the biltmore stick up to their eye and reading the corresponding number. 
 The biltmore stick can also easily be used to measure the DBH of the tree. 
-Other ways of measuring tree volume include using a combination of lasers and inclinometers or by having a person manually climb up the tree and drop a tape measure down to measure the height.
-Despite these relatively simpe ways of measuring they often leave room for error.
-That’s why a more common way of getting a precise reading is to cut down trees in the area and then physically measure them.
-After the relevant data is collected one applies a stem-profile taper equation that is specific to the species of tree measured.
-These taper functions are based on numerous measurements and have been refined over the years.
-The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions.
-This again is cost prohibitive and only so much of a forest can be accounted for in the taper function.
-Even when one manages to create a new taper function, it is still far from perfect as all trees, even from the same species, differ in shape.
+Other ways of measuring tree volume include using a combination of lasers and 
+inclinometers or by having a person manually climb up the tree and drop a tape 
+measure down to measure the height. Despite these relatively simpe ways of 
+measuring they often leave room for error. That’s why a more common way of 
+getting a precise reading is to cut down trees in the area and then physically 
+measure them. After the relevant data is collected one applies a stem-profile 
+taper equation that is specific to the species of tree measured. These taper 
+functions are based on numerous measurements and have been refined over the years. 
+The main problem with taper functions is that to get an accurate result they 
+need to be callibrated to the hundreds of local conditions. This again is cost 
+prohibitive and only so much of a forest can be accounted for in the taper function. 
+Even when one manages to create a new taper function, it is still far from 
+perfect as all trees, even from the same species, differ in shape.
 
-This project aims to solve the described problem by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation.
-This is supposed to then give a accurate result, cheaply and non-destructively.
-All within the reaches of a mobile phone.
+This project aims to solve the described problem by using a combination of 
+photos taken of the tree, various sensor data from the phone and mathematical 
+computation. This is supposed to then give a accurate result, cheaply and 
+non-destructively. All within the reaches of a mobile phone.
 
@@
 \end{figure}
-The first thing a person does when accessing the mobile application is to log into the system with their user name and password.
-After that they are prompted with a simple menu where they can either choose to add new trees to their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}.
-The data can then later be used with SilviaTerra's other products.
-If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, \ref{input}.
+The first thing a person does when accessing the mobile application is to log 
+into the system with their user name and password. After that they are 
+prompted with a simple menu where they can either choose to add new trees to 
+their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}. 
+The data can then later be used with SilviaTerra's other products. If the user 
+chooses to add a new tree then they will be asked to enter the tree's DBH and 
+species, \ref{input}. 
 \begin{figure}[!htb]
@@
 
-After that they take two pictures.
-The interface is quite simple with a cross-hair to show where to point the camera.
-The first picture is of the actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{bole}.
-Between these two pictures the camera will have to be tilted vertically to capture the bole.
-The angle between these two camera positions is recorded through the internal gyroscope.
-After this stage the user's role in the application is done and the pictures is sent to one of SilviaTerra's web servers.
-Here horizontal lines that run parallel with the bottom of the image are drawn across the entire picture.
-These are drawn at different pixel heights with equal pixel distance in between.
-The reason this is done is because computer vision is still an experimental technology and is not fully capable yet of being to detect the edges of trees with a background of forestry.
-Seeing as this is a commercial product, it would be quite disadvantageous if errors were commonplace.
-So instead the images with the newly drawn lines are then forwarded to Amazon's Mechanical Turk service.
-Just like the name suggests Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of a automated chess board.
-The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed.
-Usually the tasks are ones that are difficult for computers but easy for humans.
-SilviaTerra creates tasks where people manually click where the previously drawn lines, cross the edges of the tree.
-Ones of the points of intersection are identified, their corresponding pixel coordinates are the sent back to the web server.
-The pixel coordinates along with information from the phone is piped into a program where the volume is then calculated.
-It is this program that is the main topic of this research paper.
+After that they take two pictures. The interface is quite simple with a 
+cross-hair to show where to point the camera. The first picture is of the 
+actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, 
+figure \ref{bole}. Between these two pictures the camera will have to be tilted 
+vertically to capture the bole. The angle between these two camera positions is 
+recorded through the internal gyroscope. After this stage the user's role in 
+the application is done and the pictures is sent to one of SilviaTerra's web 
+servers. Here horizontal lines that run parallel with the bottom of the image 
+are drawn across the entire picture. These are drawn at different pixel heights 
+with equal pixel distance in between. The reason this is done is because 
+computer vision is still an experimental technology and is not fully capable 
+yet of being to detect the edges of trees with a background of forestry. 
+Seeing as this is a commercial product, it would be quite disadvantageous if 
+errors were commonplace. So instead the images with the newly drawn lines are 
+then forwarded to Amazon's Mechanical Turk service. Just like the name suggests 
+Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of 
+a automated chess board. The service is in its simplest form a crowd sourcing 
+internet marketplace where individuals or businesses can request tasks to be 
+completed. Usually the tasks are ones that are difficult for computers but 
+easy for humans. SilviaTerra creates tasks where people manually click where 
+the previously drawn lines, cross the edges of the tree. Ones of the points of 
+intersection are identified, their corresponding pixel coordinates are the sent 
+back to the web server. The pixel coordinates along with information from the 
+phone is piped into a program where the volume is then calculated. It is this 
+program that is the main topic of this research paper.
 
 \section{Method}
-Once all the required information has been retrieved it is piped into the program that will be calculating the volume.
-The dataset that is used comprises information about the phone and the results from Mechanical Turk.
-It is as follows:
+Once all the required information has been retrieved it is piped into the 
+program that will be calculating the volume. The dataset that is used comprises 
+information about the phone and the results from Mechanical Turk. It is as follows:
 \begin{itemize}
@@
 \subsection{Distance to tree}
-The DBH is used as a reference to find all the other measurements.
-The first unknown that must be discovered is the distance to the tree.
-This distance is parallel with the ground.
-The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$.
-Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels betwee them.
+The DBH is used as a reference to find all the other measurements. The first
+unknown that must be discovered is the distance to the tree. This distance is 
+parallel with the ground. The difference between the x-coordinates of the left 
+and right pixel that correspond to the DBH is calculated, figure 
+\ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of
+the DBH is now known in both pixels and in a physical measurement, the ratio
+between the two can be determined by dividing the physical distance between 
+$X_1$ and $X_2$ with the number of pixels betwee them.
 \begin{figure}[htp]
@@
 \end{figure}
- As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constructed with the horizontal angle of view as the vertex angle.
-The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree.
-To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image.
-The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, contriving the angle $\theta$ which is exactly half of the angle of view and the actual distance to the tree $d$.
-This line also divides $W_1$ into $W_2$ and futhermore producing a right angle triangle that can be solved.
-Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
+ As seen in figure \ref{horizontal_triangle} an isosceles triangle can be 
+ constructed with the horizontal angle of view as the vertex angle. The triangles 
+ two congruent legs extend out towards the base of the triangle, $W_1$, which 
+ is inline with the tree. To get the length of the base, $W_1$, the previously 
+ calculated scale between pixels and the real world is multiplied with the 
+ number of pixels in the horizontal resolution of the image. The distance to 
+ the tree can easily calculated by drawing an extra line that bisects the vertex
+ angle, contriving the angle $\theta$ which is exactly half of the angle of view 
+ and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ 
+ and futhermore producing a right angle triangle that can be solved. Finally to 
+ actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
 
@@
 
-The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree.
-The first step taken is to get the height.
-The vertical image plate scale which is the ratio between degrees and pixels in the vertical axis is calculated by dividing the vertical resolution of the image with the vertical angle of view.
-This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane.
-With this method it is possible to find out the angle to each pixel coordinate, $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
-This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}.
-However if the pixel coordinate was identified in the second image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}.
-The diagonal distance to $P$ is also calculated so that the width can later be determined.
+The next stage is retrieving the real world height to each set pixel coordinates 
+as well as the width between them, which will be the diameter of the tree. The 
+first step taken is to get the height. The vertical image plate scale which is 
+the ratio between degrees and pixels in the vertical axis is calculated by 
+dividing the vertical resolution of the image with the vertical angle of view. 
+This can be done because it is assumed that the silhouette of the tree is on a
+plane in the real world which corresponds to the image plane. With this method 
+it is possible to find out the angle to each pixel coordinate, $P$ in figure 
+\ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is 
+denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However 
+if the pixel coordinate was identified in the second image, when the phone was 
+tilted upwards, then angle between the distance to the tree and the lower boundary
+of the vertical angle of view, $B$, will need to be added, which is represented
+by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also 
+calculated so that the width can later be determined.
 
-The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them.
-This is accomplished by employing the method for calculating the distance to the tree, in reverse.
-In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
-$W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$.
-$W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
-To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.
+The final part of obataining the parameters for the set of pixel coordinates at 
+that height is to calculate the width between them. This is accomplished by 
+employing the method for calculating the distance to the tree, in reverse. 
+In this case what is known in figure \ref{horizontal_triangle} is the distance 
+$d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ 
+is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again 
+multiplied by 2 and then divided by the number of pixels in the horizontal 
+resolution of the image, once more returning the ratio between pixels and the 
+corresponding real world measurement. To finish off this ratio is the multiplied
+by the number of pixels between $X_1$ and $X_2$ dispensing the physical width
+of the tree at that certain height.
 
 \subsection{Volume}
-Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume.
-The current implemented method of doing this is by integrating all the points using the trapezoidal rule.
-All the widths are divided by 2 to get the radius.
-Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height.
-The basal area is plotted along with their corresponding heights.
-The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
+Now the only thing left to produce the final result is to use these newly 
+calculated widths and heights to obtain the total volume. The current implemented 
+method of doing this is by integrating all the points using the trapezoidal rule.
+All the widths are divided by 2 to get the radius. Then they are squared and
+multiplied with $\pi$ to the obtain the basal area at each height. The basal 
+area is plotted along with their corresponding heights. The trapezoidal rule 
+then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
 The area between all the points is the summed together.
 
-This whole process is looped through for every identified pixel coordinate set until the width and height is known for all of them.
-Once that is done, a new loop is created for calculating the total volume.
+This whole process is looped through for every identified pixel coordinate set 
+until the width and height is known for all of them. Once that is done, a new
+loop is created for calculating the total volume.
 
 \subsection{Simulation}
-To test this method a \emph{virtual tree} was constructed.
-This tree was made out of blue tape and tapered off so that the program could be thoroughly tested.
-Along the tape points were marked out and the height to them plus the width between corresponding points were measured.
-Just like in the real application, two pictures were taken.
-One of the DBH and one of the top of the tree.
-Each set of points was then identified on the pictures and measured in pixels.
-The parameters were then passed into the program and the results were cross-referenced with the real life measurements.
+To test this method a \emph{virtual tree} was constructed. This tree was made 
+out of blue tape and tapered off so that the program could be thoroughly tested. 
+Along the tape points were marked out and the height to them plus the width
+between corresponding points were measured. Just like in the real application, 
+two pictures were taken. One of the DBH and one of the top of the tree. Each 
+set of points was then identified on the pictures and measured in pixels. The 
+parameters were then passed into the program and the results were cross-referenced
+with the real life measurements.
 

### 2013-07-06 paper.tex
Added discussion section with error tolerance subsection

@@
 \end{table}
+
+\section{Discussion}
+\subsection{Error tolerance}
+As seen in table \ref{heights} either when the vertical tilt angle is very wide, 
+standing close to the tree, or very narrow, standing far away, leads to error. I
+t's not clear what is causing this but there a two possibl reasons. The first one is
+that because of the short focal length of the camera and overall poor quality when 
+compared to a real camera, causes various forms of distrotion, especially around the 
+edges of the image. When the person is standing close to the tree so the vertical tilt 
+angle is over 45\degree, the points tend to lie exactly there, around the edges of 
+the image. The distortion then affects these points causing the pixel measurements 
+to be incorrect. The other problem occurs in the direct opposite situation, when 
+the person is standing far away and the angle is very narrow. Naturally what happens 
+is that the resolution of the camera just isn't high enough and the pixel measurements 
+just aren't accurate anymore. However most of this distortion seems to only affect the 
+vertical measurements because when looking at table \ref{widths} it is clearly seen that 
+the readings do not stray much at all from their actual values. Despite this the deviation 
+in the height readings still alters the volume calculations.
+\subsection{Limitations}
+\subsection{Future development}
+\subsection{Future development}

### 2013-07-06 paper.tex
Spell checking, rewording and added future development + limitations section

--- 2013-07-06 paper.tex
@@
 \section{Introduction}
-Tree volume measurement is one of the many parameters in documenting and 
-profiling trees and is at the center of forest inventory. The use of this 
-information is important to companies who need to make smart, profitable 
-decisions but also need to be certain that the choices they make now are 
-maintable in the future. However this information also benefits climate 
-modellers, environmental groups and governments who want to be able to sustain 
-ecosystems and plan restorations. To be able to accomplish this in forestry, 
-it is a necessity to be able to collect large amounts of data easily and on a 
-regular basis. Most methods used today to measure individual trees are often 
-done manually with a variety of tools. The use of these analogue approaches 
-is expensive, time consuming and in most cases the data collected doesn't meet 
-today's standards \cite{digital imaged based tree measurement for forest inventory}. 
-SilviaTerra, however, is a company willing to change that and have already 
-developed several tools for scanning, analyzing and mapping areas of forestry 
-to give accurate results. This is achieved through computer exactitude combined 
-with statistical analysis. Recently a new project was started, with this 
-research paper developing the mathematical back end, to allow users to easily 
-get a correct volume measurement by only using their smart phone.
+Tree volume measurement is one of the many parameters in documenting and profili
+ng trees and is at the center of forest inventory. The use of this information i
+s important to companies who need to make smart, profitable decisions but also n
+eed to be certain that the choices they make now are maintable in the future. Ho
+wever this information also benefits climate modellers, environmental groups and
+ governments who want to be able to sustain ecosystems and plan restorations. To
+ be able to accomplish this in forestry, it is a necessity to be able to collect
+ large amounts of data easily and on a regular basis. Most methods used today to
+ measure individual trees are often done manually with a variety of tools. The u
+se of these analogue approaches is expensive, time consuming and in most cases t
+he data collected doesn't meet today's standards \cite{digital imaged based tree
+ measurement for forest inventory}. SilviaTerra, however, is a company willing t
+o change that and have already developed several tools for scanning, analyzing a
+nd mapping areas of forestry to give accurate results. This is achieved through 
+computer exactitude combined with statistical analysis. Recently a new project w
+as started, with this research project developing the mathematical back end, to 
+allow users to easily get a correct volume measurement by only using their smart
+ phone.
 
 The measurements required for forest inventory can be achieved through a number 
-of methods. One of them is using a so called biltmore stick. A biltmore stick 
-is similar to a normal ruler but has been specially modified for measuring trees. 
+of methods. One of them is using a so called biltmore stick. A biltmore stick is
+ similar to a normal ruler but has been specially modified for measuring trees. 
 When standing at a certain distance the user can discern the height of the tree 
-by holding the biltmore stick up to their eye and reading the corresponding number. 
-The biltmore stick can also easily be used to measure the DBH of the tree. 
-Other ways of measuring tree volume include using a combination of lasers and 
-inclinometers or by having a person manually climb up the tree and drop a tape 
-measure down to measure the height. Despite these relatively simpe ways of 
-measuring they often leave room for error. That’s why a more common way of 
-getting a precise reading is to cut down trees in the area and then physically 
-measure them. After the relevant data is collected one applies a stem-profile 
-taper equation that is specific to the species of tree measured. These taper 
-functions are based on numerous measurements and have been refined over the years. 
-The main problem with taper functions is that to get an accurate result they 
-need to be callibrated to the hundreds of local conditions. This again is cost 
-prohibitive and only so much of a forest can be accounted for in the taper function. 
-Even when one manages to create a new taper function, it is still far from 
-perfect as all trees, even from the same species, differ in shape.
+by holding the biltmore stick up to their eye and reading the corresponding numb
+er. The biltmore stick can also easily be used to measure the DBH of the tree. O
+ther ways of measuring tree volume include using a combination of lasers and inc
+linometers or by having a person manually climb up the tree and drop a tape meas
+ure down to measure the height. Despite these relatively simple ways of measurin
+g they often leave room for error. That’s why a more common way of getting a pre
+cise reading is to cut down trees in the area and then physically measuring them
+. After the relevant data is collected one applies a stem-profile taper equation
+ that is specific to the species of tree measured. These taper functions are bas
+ed on numerous measurements and have been refined over the years. The main probl
+em with taper functions is that to get an accurate result they need to be callib
+rated to the hundreds of local conditions. This again is cost prohibitive and on
+ly so much of a forest can be accounted for in the taper function. Even when one
+ manages to create a new taper function, it is still far from perfect as all tre
+es, even from the same species, differ in shape.
 
-This project aims to solve the described problem by using a combination of 
-photos taken of the tree, various sensor data from the phone and mathematical 
-computation. This is supposed to then give a accurate result, cheaply and 
-non-destructively. All within the reaches of a mobile phone.
+This project aims to solve the described problem by using a combination of photo
+s taken of the tree, various sensor data from the phone and mathematical computa
+tion. This is supposed to then give a accurate result, cheaply and non-destructi
+vely. All within the reaches of a mobile phone.
 
@@
 \end{figure}
-The first thing a person does when accessing the mobile application is to log 
-into the system with their user name and password. After that they are 
-prompted with a simple menu where they can either choose to add new trees to 
-their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}. 
-The data can then later be used with SilviaTerra's other products. If the user 
-chooses to add a new tree then they will be asked to enter the tree's DBH and 
-species, \ref{input}. 
+The first thing a person does when accessing the mobile application is to log in
+to the system with their user name and password. After that they are prompted wi
+th a simple menu where they can either choose to add new trees to their inventor
+y or upload them to SilviaTerra's cloud service, figure \ref{main}. The data can
+ then later be used with SilviaTerra's other products. If the user chooses to ad
+d a new tree then they will be asked to enter the tree's DBH and species, \ref{i
+nput}. 
 \begin{figure}[!htb]
@@
 
-After that they take two pictures. The interface is quite simple with a 
-cross-hair to show where to point the camera. The first picture is of the 
-actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, 
-figure \ref{bole}. Between these two pictures the camera will have to be tilted 
-vertically to capture the bole. The angle between these two camera positions is 
-recorded through the internal gyroscope. After this stage the user's role in 
-the application is done and the pictures is sent to one of SilviaTerra's web 
-servers. Here horizontal lines that run parallel with the bottom of the image 
-are drawn across the entire picture. These are drawn at different pixel heights 
-with equal pixel distance in between. The reason this is done is because 
-computer vision is still an experimental technology and is not fully capable 
-yet of being to detect the edges of trees with a background of forestry. 
-Seeing as this is a commercial product, it would be quite disadvantageous if 
-errors were commonplace. So instead the images with the newly drawn lines are 
-then forwarded to Amazon's Mechanical Turk service. Just like the name suggests 
-Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of 
-a automated chess board. The service is in its simplest form a crowd sourcing 
-internet marketplace where individuals or businesses can request tasks to be 
-completed. Usually the tasks are ones that are difficult for computers but 
-easy for humans. SilviaTerra creates tasks where people manually click where 
-the previously drawn lines, cross the edges of the tree. Ones of the points of 
-intersection are identified, their corresponding pixel coordinates are the sent 
-back to the web server. The pixel coordinates along with information from the 
-phone is piped into a program where the volume is then calculated. It is this 
-program that is the main topic of this research paper.
+After that they take two pictures. The interface is quite simple with a cross-ha
+ir to show where to point the camera. The first picture is of the actual DBH, fi
+gure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{b
+ole}. Between these two pictures the camera will have to be tilted vertically to
+ capture the bole. The angle between these two camera positions is recorded thro
+ugh the internal gyroscope. After this stage the user's role in the application 
+is done and the pictures is sent to one of SilviaTerra's web servers. Here horiz
+ontal lines that run parallel with the bottom of the image are drawn across the 
+entire picture. These are drawn at different pixel heights with equal pixel dist
+ance in between. The reason this is done is because computer vision is still an 
+experimental technology and is not fully capable yet of being to detect the edge
+s of trees with a background of forestry. Seeing as this is a commercial product
+, it would be quite disadvantageous if errors were commonplace. So instead the i
+mages with the newly drawn lines are then forwarded to Amazon's Mechanical Turk 
+service. Just like the name suggests Amazon's Mechanical Turk is meant to be an 
+analogy to the 18th century concept of a automated chess board. The service is i
+n its simplest form a crowd sourcing internet marketplace where individuals or b
+usinesses can request tasks to be completed. Usually the tasks are ones that are
+ difficult for computers but easy for humans. SilviaTerra creates tasks where pe
+ople manually click where the previously drawn lines, cross the edges of the tre
+e. Once the points of intersection are identified, their corresponding pixel coo
+rdinates are the sent back to the web server. The pixel coordinates along with i
+nformation from the phone is piped into a program where the volume is then calcu
+lated. It is this program that is the main topic of this research paper.
 
 \section{Method}
-Once all the required information has been retrieved it is piped into the 
-program that will be calculating the volume. The dataset that is used comprises 
-information about the phone and the results from Mechanical Turk. It is as follows:
+Once all the required information has been retrieved it is piped into the progra
+m that will be calculating the volume. The dataset that is used comprises inform
+ation about the phone and the results from Mechanical Turk. It is as follows:
 \begin{itemize}
@@
 \subsection{Distance to tree}
-The DBH is used as a reference to find all the other measurements. The first
-unknown that must be discovered is the distance to the tree. This distance is 
-parallel with the ground. The difference between the x-coordinates of the left 
-and right pixel that correspond to the DBH is calculated, figure 
-\ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of
-the DBH is now known in both pixels and in a physical measurement, the ratio
-between the two can be determined by dividing the physical distance between 
-$X_1$ and $X_2$ with the number of pixels betwee them.
+The DBH is used as a reference to find all the other measurements. The first unk
+nown that must be discovered is the distance to the tree. This distance is paral
+lel with the ground. The difference between the x-coordinates of the left and ri
+ght pixel that correspond to the DBH is calculated, figure \ref{horizontal_trian
+gle}, $X_1$ and $X_2$. Due to the fact that the width of the DBH is now known in
+ both pixels and in a physical measurement, the ratio between the two can be det
+ermined by dividing the physical distance between $X_1$ and $X_2$ with the numbe
+r of pixels betwee them.
 \begin{figure}[htp]
@@
 \end{figure}
- As seen in figure \ref{horizontal_triangle} an isosceles triangle can be 
- constructed with the horizontal angle of view as the vertex angle. The triangles 
- two congruent legs extend out towards the base of the triangle, $W_1$, which 
- is inline with the tree. To get the length of the base, $W_1$, the previously 
- calculated scale between pixels and the real world is multiplied with the 
- number of pixels in the horizontal resolution of the image. The distance to 
- the tree can easily calculated by drawing an extra line that bisects the vertex
- angle, contriving the angle $\theta$ which is exactly half of the angle of view 
- and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ 
- and futhermore producing a right angle triangle that can be solved. Finally to 
- actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
+ As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constr
+ucted with the horizontal angle of view as the vertex angle. The triangles two c
+ongruent legs extend out towards the base of the triangle, $W_1$, which is inlin
+e with the tree. To get the length of the base, $W_1$, the previously calculated
+ scale between pixels and the real world is multiplied with the number of pixels
+ in the horizontal resolution of the image. The distance to the tree can easily 
+calculated by drawing an extra line that bisects the vertex angle, contriving th
+e angle $\theta$ which is exactly half of the angle of view and the actual dista
+nce to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore prod
+ucing a right angle triangle that can be solved. Finally to actually calculate t
+he distance, $d$, $W_2$ is divided by $\tan{\theta}$.
 
@@
  as well as the width between them, which will be the diameter of the tree. The 
-first step taken is to get the height. The vertical image plate scale which is 
-the ratio between degrees and pixels in the vertical axis is calculated by 
-dividing the vertical resolution of the image with the vertical angle of view. 
-This can be done because it is assumed that the silhouette of the tree is on a
-plane in the real world which corresponds to the image plane. With this method 
-it is possible to find out the angle to each pixel coordinate, $P$ in figure 
-\ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is 
-denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However 
-if the pixel coordinate was identified in the second image, when the phone was 
-tilted upwards, then angle between the distance to the tree and the lower boundary
-of the vertical angle of view, $B$, will need to be added, which is represented
-by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also 
-calculated so that the width can later be determined.
+first step taken is to get the height. The vertical image plate scale, which is 
+the ratio between degrees and pixels in the vertical axis, is calculated by divi
+ding the vertical resolution of the image with the vertical angle of view. This 
+can be done because it is assumed that the silhouette of the tree is on a plane 
+in the real world which corresponds to the image plane on the camera sensor. Wit
+h this method it is possible to find out the angle to each pixel coordinate, exa
+mple $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. Th
+is is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{trian
+gle2}. However if the pixel coordinate was identified in the bole image, when th
+e phone was tilted upwards, then angle between the distance to the tree and the 
+lower boundary of the vertical angle of view, $B$, will need to be added, which 
+is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to 
+$P$ is also calculated so that the width can later be determined.
 
 The final part of obataining the parameters for the set of pixel coordinates at 
-that height is to calculate the width between them. This is accomplished by 
-employing the method for calculating the distance to the tree, in reverse. 
-In this case what is known in figure \ref{horizontal_triangle} is the distance 
-$d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ 
-is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again 
-multiplied by 2 and then divided by the number of pixels in the horizontal 
-resolution of the image, once more returning the ratio between pixels and the 
-corresponding real world measurement. To finish off this ratio is the multiplied
-by the number of pixels between $X_1$ and $X_2$ dispensing the physical width
-of the tree at that certain height.
+that height is to calculate the width between them. This is accomplished by empl
+oying the method for calculating the distance to the tree, in reverse. In this c
+ase what is known in figure \ref{horizontal_triangle} is the distance $d$ and th
+e angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ is calcula
+ted by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again multiplied by 2
+ and then divided by the number of pixels in the horizontal resolution of the im
+age, once more returning the ratio between pixels and the corresponding real wor
+ld measurement. To finish off this ratio is the multiplied by the number of pixe
+ls between $X_1$ and $X_2$ dispensing the physical width of the tree at that cer
+tain height.
 
 \subsection{Volume}
-Now the only thing left to produce the final result is to use these newly 
-calculated widths and heights to obtain the total volume. The current implemented 
-method of doing this is by integrating all the points using the trapezoidal rule.
-All the widths are divided by 2 to get the radius. Then they are squared and
-multiplied with $\pi$ to the obtain the basal area at each height. The basal 
-area is plotted along with their corresponding heights. The trapezoidal rule 
-then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
-The area between all the points is the summed together.
+Now the only thing left to produce the final result is to use these newly calcul
+ated widths and heights to obtain the total volume. The current implemented meth
+od of doing this is by integrating all the points using the trapezoidal rule. Al
+l the widths are divided by 2 to get the radius. Then they are squared and multi
+plied with $\pi$ to the obtain the basal area at each height. The basal area is 
+plotted along with their corresponding heights. The trapezoidal rule then define
+s that the area between two points is: $\frac{a_1 + a_2}{2} \times h$. The area 
+between all the points is then summed together.
 
-This whole process is looped through for every identified pixel coordinate set 
-until the width and height is known for all of them. Once that is done, a new
-loop is created for calculating the total volume.
+This whole process is looped through for every identified pixel coordinate set u
+ntil the width and height is known for all of them. Once that is done, a new loo
+p is created for calculating the total volume.
 
 \subsection{Simulation}
-To test this method a \emph{virtual tree} was constructed. This tree was made 
-out of blue tape and tapered off so that the program could be thoroughly tested. 
-Along the tape points were marked out and the height to them plus the width
-between corresponding points were measured. Just like in the real application, 
-two pictures were taken. One of the DBH and one of the top of the tree. Each 
-set of points was then identified on the pictures and measured in pixels. The 
-parameters were then passed into the program and the results were cross-referenced
-with the real life measurements.
+To test this method a \emph{virtual tree}, that represented a 4m tall tree,was c
+onstructed. This tree was made out of blue tape and tapered off drastically so t
+hat the program could be thoroughly tested. Along the tape, points were marked o
+ut and the height to them plus the width between corresponding points was measur
+ed. Just like in the real application, two pictures were taken. One of the DBH a
+nd one of the top of the tree. Each set of points was then identified on the pic
+tures and measured in pixels. The parameters were then passed into the program a
+nd the results were cross-referenced with the real life measurements.
 
@@
 		\hline
-		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
+		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degr
+ee & 15\degree \\
 		\hline
@@
 		\hline
-		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
+		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degr
+ee & 15\degree \\
 		\hline
-		12       & 11.7      & 12.1      & 12.2      & 12.1      & 12.3      & 12.6      & 11.9      \\
-		27       & 26.8      & 28.0      & 28.4      & 28.2      & 28.5      & 28.4      & 28.1      \\
-		37       & 36.4      & 37.5      & 38.0      & 37.7      & 38.2      & 38.1      & 37.7      \\
-		44       & 42.4      & 43.6      & 44.4      & 44.0      & 44.4      & 44.0      & 44.4      \\
-		52       & 51.3      & 52.7      & 53.5      & 53.0      & 53.4      & 52.9      & 53.2      \\
-		61       & 60.9      & 62.2      & 63.1      & 62.1      & 62.8      & 62.5      & 61.1      \\
-		59       & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      \\
+		12       & 11.7      & 12.1      & 12.2      & 12.1      & 12.3      & 12.6   
+   & 11.9      \\
+		27       & 26.8      & 28.0      & 28.4      & 28.2      & 28.5      & 28.4   
+   & 28.1      \\
+		37       & 36.4      & 37.5      & 38.0      & 37.7      & 38.2      & 38.1   
+   & 37.7      \\
+		44       & 42.4      & 43.6      & 44.4      & 44.0      & 44.4      & 44.0   
+   & 44.4      \\
+		52       & 51.3      & 52.7      & 53.5      & 53.0      & 53.4      & 52.9   
+   & 53.2      \\
+		61       & 60.9      & 62.2      & 63.1      & 62.1      & 62.8      & 62.5   
+   & 61.1      \\
+		59       & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0   
+   & 59.0      \\
 		\hline
@@
     	\hline
-		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
-		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401    & 445624    \\
+		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degr
+ee & 15\degree \\
+		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401 
+   & 445624    \\
 		\hline
@@
 \end{table}
+
+\section{Discussion}
+\subsection{Error tolerance}
+As seen in table \ref{heights} either when the vertical tilt angle is very wide,
+ standing close to the tree, or very narrow, standing far away, leads to error. 
+It's not clear what is causing this but there a two possibl reasons. The first o
+ne is that because of the short focal length of the camera and overall poor qual
+ity when compared to a real camera, causes various forms of distrotion, especial
+ly around the edges of the image. When the person is standing close to the tree 
+so the vertical tilt angle is over 45\degree, the points tend to lie exactly the
+re, around the edges of the image. The distortion then affects these points caus
+ing the pixel measurements to be incorrect. The other problem occurs in the dire
+ct opposite situation, when the person is standing far away and the angle is ver
+y narrow. Naturally what happens is that the resolution of the camera just isn't
+ high enough and the pixel measurements just aren't accurate anymore. However mo
+st of this distortion seems to only affect the vertical measurements because whe
+n looking at table \ref{widths} it is clearly seen that the readings do not stra
+y much at all from their actual values. Despite this the deviation in the height
+ readings still alters the volume calculations. Ideally a person should stand so
+ that the vertical tilt angle lies between 45\degree and 25\degree which is basi
+cally means they should stand somewhere between one tree length to two tree leng
+th's from the tree.
+\subsection{Limitations}
+As pointed out in the previous section the main hindrance attatched to this appl
+ication is the actual camera. 
+\subsection{Future development}
+The main area that could be expanded on is the edge detection. As explained in t
+he background, today computer vision is still far from perfect. However that wil
+l hopefully change in the near future and things like Mechanical Turk can be dro
+pped for a more efficient system that produces little to no error and works auto
+nomously.

### 2013-07-06 paper.tex

@@
 Tree volume measurement is one of the many parameters in documenting and profili
-ng trees and is at the center of forest inventory. The use of this information i
+ng trees and is at the center of forest inventory. 
+% Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration
+
+%I'd delete this part
+The use of this information i
 s important to companies who need to make smart, profitable decisions but also n
@@
 wever this information also benefits climate modellers, environmental groups and
- governments who want to be able to sustain ecosystems and plan restorations. To
+ governments who want to be able to sustain ecosystems and plan restorations.
+
+ %Unfortunately, estimating tree volume is currently a complex, time-consuming, and expensive procedure.
+ To
  be able to accomplish this in forestry, it is a necessity to be able to collect
@@
 he data collected doesn't meet today's standards \cite{digital imaged based tree
- measurement for forest inventory}. SilviaTerra, however, is a company willing t
-o change that and have already developed several tools for scanning, analyzing a
-nd mapping areas of forestry to give accurate results. This is achieved through 
-computer exactitude combined with statistical analysis. Recently a new project w
-as started, with this research project developing the mathematical back end, to 
-allow users to easily get a correct volume measurement by only using their smart
- phone.
+ measurement for forest inventory}.
 
 The measurements required for forest inventory can be achieved through a number 
-of methods. One of them is using a so called biltmore stick. A biltmore stick is
+of methods. %the most common method uses what is called a biltmore stick...
+One of them is using a so called biltmore stick. A biltmore stick is
  similar to a normal ruler but has been specially modified for measuring trees. 
@@
 by holding the biltmore stick up to their eye and reading the corresponding numb
-er. The biltmore stick can also easily be used to measure the DBH of the tree. O
-ther ways of measuring tree volume include using a combination of lasers and inc
+er. The biltmore stick can also easily be used to measure the DBH of the tree. %define DBH
+
+%i'd eliminate this section
+Other ways of measuring tree volume include using a combination of lasers and inc
 linometers or by having a person manually climb up the tree and drop a tape meas
-ure down to measure the height. Despite these relatively simple ways of measurin
+ure down to measure the height.
+
+Despite these relatively simple ways of measurin
 g they often leave room for error. That’s why a more common way of getting a pre
-cise reading is to cut down trees in the area and then physically measuring them
-. After the relevant data is collected one applies a stem-profile taper equation
- that is specific to the species of tree measured. These taper functions are bas
+cise reading is to cut down trees in the area and then physically measure them
+. After the relevant data is collected, one applies a stem-profile taper equation
+ that is specific to the species of tree measured. These taper functions are bas %need more explanation about taper functions
 ed on numerous measurements and have been refined over the years. The main probl
@@
 
-This project aims to solve the described problem by using a combination of photo
+%However, a high-tech startup company named SilviaTerra is trying to change that.
+ SilviaTerra, however, is a company willing t
+o change that and have already developed several tools for scanning, analyzing a
+nd mapping areas of forestry to give accurate results. This is achieved through 
+computer exactitude combined with statistical analysis. Recently a new project w %this part is awkwardly phrased
+as started, with this research project developing the mathematical back end, to 
+allow users to easily get a correct volume measurement by only using their smart
+ phone.
+
+This project aims to solve 
+the described problem % -> tree volumes by...
+using a combination of photo
 s taken of the tree, various sensor data from the phone and mathematical computa
 tion. This is supposed to then give a accurate result, cheaply and non-destructi
-vely. All within the reaches of a mobile phone.
+vely. All within the reaches of a mobile phone. % this last sentence is awkward
 
@@
 \end{figure}
+%not sure you need this part... just stick to describing the bare-minimum needed to understand the measurements
 The first thing a person does when accessing the mobile application is to log in
@@
 th a simple menu where they can either choose to add new trees to their inventor
-y or upload them to SilviaTerra's cloud service, figure \ref{main}. The data can
- then later be used with SilviaTerra's other products. If the user chooses to ad
+y or upload them to SilviaTerra's cloud service, figure \ref{main}. 
+The data can then later be used with SilviaTerra's other products.
+
+ If the user chooses to ad
 d a new tree then they will be asked to enter the tree's DBH and species, \ref{i
@@
  capture the bole. The angle between these two camera positions is recorded thro
-ugh the internal gyroscope. After this stage the user's role in the application 
-is done and the pictures is sent to one of SilviaTerra's web servers. Here horiz
+ugh the internal gyroscope. %and accelerometers
+
+%explain that, "Now the challenge is to extract useful measurements from these pictures"
+% "We need to identify a bunch of points along the trunk of the tree to fit taper equations"
+After this stage the user's role in the application 
+is done and the pictures are sent to one of SilviaTerra's web servers. Here horiz
 ontal lines that run parallel with the bottom of the image are drawn across the 
 entire picture. These are drawn at different pixel heights with equal pixel dist
-ance in between. The reason this is done is because computer vision is still an 
+ance in between.
+
+%Say instead, "to perform this task computationally would be quite difficult and error-prone,
+% so instead the company opted to have this task performed manually by workers on Amazon's Mechanical Turk service.
+The reason this is done is because computer vision is still an 
 experimental technology and is not fully capable yet of being to detect the edge
-s of trees with a background of forestry. Seeing as this is a commercial product
-, it would be quite disadvantageous if errors were commonplace. So instead the i
+s of trees with a background of forestry.
+
+%eliminate this part
+Seeing as this is a commercial product
+, it would be quite disadvantageous if errors were commonplace.
+
+%not sure you need this part - maintain focus on the key problem of solving volumes!
+So instead the i
 mages with the newly drawn lines are then forwarded to Amazon's Mechanical Turk 
@@
 usinesses can request tasks to be completed. Usually the tasks are ones that are
- difficult for computers but easy for humans. SilviaTerra creates tasks where pe
+ difficult for computers but easy for humans.
+
+ SilviaTerra creates tasks where pe
 ople manually click where the previously drawn lines, cross the edges of the tre
 e. Once the points of intersection are identified, their corresponding pixel coo
-rdinates are the sent back to the web server. The pixel coordinates along with i
+rdinates are the sent back to the web server.
+
+%The pixel coordinates that result are then paired with the phone sensor data 
+% and input into a program to perform a photogrammetric analysis to derive the taper equations.
+The pixel coordinates along with i
 nformation from the phone is piped into a program where the volume is then calcu
@@
 Once all the required information has been retrieved it is piped into the progra
-m that will be calculating the volume. The dataset that is used comprises inform
+m that will be calculating the volume. %and taper equations!
+
+The dataset that is used comprises inform
 ation about the phone and the results from Mechanical Turk. It is as follows:
@@
 \subsection{Distance to tree}
+%Try to make this section clearer - it's a bit tedious to understand
 The DBH is used as a reference to find all the other measurements. The first unk
@@
 ongruent legs extend out towards the base of the triangle, $W_1$, which is inlin
-e with the tree. To get the length of the base, $W_1$, the previously calculated
+e with the tree. To get the length of the base, %(corresponding to <WHAT> in the real world...)
+$W_1$, the previously calculated
  scale between pixels and the real world is multiplied with the number of pixels
  in the horizontal resolution of the image. The distance to the tree can easily 
-calculated by drawing an extra line that bisects the vertex angle, contriving th
+calculated by drawing an extra line that bisects the vertex angle, contriving th %contriving?  how about "resulting in an"
 e angle $\theta$ which is exactly half of the angle of view and the actual dista
-nce to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore prod
+nce to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore prod %produces* - parallel tenses
 ucing a right angle triangle that can be solved. Finally to actually calculate t
@@
 The next stage is retrieving the real world height to each set pixel coordinates
- as well as the width between them, which will be the diameter of the tree. The 
-first step taken is to get the height. The vertical image plate scale, which is 
+ as well as the width between them, which will be the diameter of the tree. %at that height
+ The 
+first step taken is to get the height. The vertical image plate scale, which is %plate scale?  What is an image plate scale?
 the ratio between degrees and pixels in the vertical axis, is calculated by divi
 ding the vertical resolution of the image with the vertical angle of view. This 
-can be done because it is assumed that the silhouette of the tree is on a plane 
+can be done because it is assumed that the silhouette of the tree is on a plane %does this just mean that the tree is straight and doesn't curve towards or away from the image?
 in the real world which corresponds to the image plane on the camera sensor. Wit
@@
 
-The final part of obataining the parameters for the set of pixel coordinates at 
+The final part of obataining %spelling!
+the parameters for the set of pixel coordinates at 
 that height is to calculate the width between them. This is accomplished by empl
@@
 ld measurement. To finish off this ratio is the multiplied by the number of pixe
-ls between $X_1$ and $X_2$ dispensing the physical width of the tree at that cer
+ls between $X_1$ and $X_2$ dispensing %resulting in
+the physical width of the tree at that cer
 tain height.
@@
 nd the results were cross-referenced with the real life measurements.
+
+%do you have a picture?
 
@@
  standing close to the tree, or very narrow, standing far away, leads to error. 
-It's not clear what is causing this but there a two possibl reasons. The first o
+It's not clear what is causing this but there a two possible reasons. The first o
 ne is that because of the short focal length of the camera and overall poor qual
-ity when compared to a real camera, causes various forms of distrotion, especial
+ity when compared to a real camera, causes various forms of distortion, especial
 ly around the edges of the image. When the person is standing close to the tree 

### 2013-07-07 paper.tex
Updated and reworked parts according to the prvious comments.

--- 2013-07-06 paper.tex
@@
 \section{Introduction}
-Tree volume measurement is one of the many parameters in documenting and 
-profiling trees and is at the center of forest inventory. The use of this 
-information is important to companies who need to make smart, profitable 
-decisions but also need to be certain that the choices they make now are 
-maintable in the future. However this information also benefits climate 
-modellers, environmental groups and governments who want to be able to sustain 
-ecosystems and plan restorations. To be able to accomplish this in forestry, 
-it is a necessity to be able to collect large amounts of data easily and on a 
-regular basis. Most methods used today to measure individual trees are often 
-done manually with a variety of tools. The use of these analogue approaches 
-is expensive, time consuming and in most cases the data collected doesn't meet 
-today's standards \cite{digital imaged based tree measurement for forest inventory}. 
-SilviaTerra, however, is a company willing to change that and have already 
-developed several tools for scanning, analyzing and mapping areas of forestry 
-to give accurate results. This is achieved through computer exactitude combined 
-with statistical analysis. Recently a new project was started, with this 
-research paper developing the mathematical back end, to allow users to easily 
-get a correct volume measurement by only using their smart phone.
+Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration.
+The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times.
+To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.
 
-The measurements required for forest inventory can be achieved through a number 
-of methods. One of them is using a so called biltmore stick. A biltmore stick 
-is similar to a normal ruler but has been specially modified for measuring trees. 
-When standing at a certain distance the user can discern the height of the tree 
-by holding the biltmore stick up to their eye and reading the corresponding number. 
-The biltmore stick can also easily be used to measure the DBH of the tree. 
-Other ways of measuring tree volume include using a combination of lasers and 
-inclinometers or by having a person manually climb up the tree and drop a tape 
-measure down to measure the height. Despite these relatively simpe ways of 
-measuring they often leave room for error. That’s why a more common way of 
-getting a precise reading is to cut down trees in the area and then physically 
-measure them. After the relevant data is collected one applies a stem-profile 
-taper equation that is specific to the species of tree measured. These taper 
-functions are based on numerous measurements and have been refined over the years. 
-The main problem with taper functions is that to get an accurate result they 
-need to be callibrated to the hundreds of local conditions. This again is cost 
-prohibitive and only so much of a forest can be accounted for in the taper function. 
-Even when one manages to create a new taper function, it is still far from 
-perfect as all trees, even from the same species, differ in shape.
+Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
+Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches do not come without their downfalls.
 
-This project aims to solve the described problem by using a combination of 
-photos taken of the tree, various sensor data from the phone and mathematical 
-computation. This is supposed to then give a accurate result, cheaply and 
-non-destructively. All within the reaches of a mobile phone.
+The measurements required for forest inventory can be achieved through a number of methods.
+One of them is using a so called biltmore stick.
+A biltmore stick is similar to a normal ruler but has been specially modified for measuring trees.
+When standing at a certain distance the user can discern the height of the tree by holding the biltmore stick up to their eye and reading the corresponding number.
+The biltmore stick can also easily be used to measure the \emph{DBH} (Diameter at Breast Height) of the tree.
+Of course a biltmore stick tries to be as generic as possible so that it can be used on a variety of trees and therefore has no guarantee of providing an accurate result.
+Other ways of measuring tree volume include using a combination of lasers and inclinometers or by having a person manually climb up the tree and drop a tape measure down to measure the height.
+Despite these relatively simple ways of measuring they often leave room for error.
+That’s why a more common way of getting a precise reading is to cut down trees in the area and then physically measuring them.
+After the relevant data is collected one applies a stem-profile taper equation that is specific to the species of tree measured.
+These taper functions are based on numerous measurements and have been refined over the years.
+The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions and are usually custom made for a certain species in a certain area.
+This again is cost prohibitive and only so much of a forest can be accounted for in the taper function.
+Even when one manages to create a new taper function, it is still far from perfect as all trees, even from the same species, differ in shape.
+
+SilviaTerra, a high-tech startup company, however is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results.
+This is achieved through computer exactitude combined with statistical analysis.
+Recently a new project was started, with this research project developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.This project aims to solve the problem of obtaining an accurate volume reading of tree, cheaply and non-destructively.
+This will be done by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation.
 
@@
 \end{figure}
-The first thing a person does when accessing the mobile application is to log 
-into the system with their user name and password. After that they are 
-prompted with a simple menu where they can either choose to add new trees to 
-their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}. 
-The data can then later be used with SilviaTerra's other products. If the user 
-chooses to add a new tree then they will be asked to enter the tree's DBH and 
-species, \ref{input}. 
+The first thing a person does when accessing the mobile application is to log into the system with their user name and password.
+After that they are prompted with a simple menu where they can either choose to add new trees to their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}.
+The data can then later be used with SilviaTerra's other products.
+If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, \ref{input}.
 \begin{figure}[!htb]
@@
 
-After that they take two pictures. The interface is quite simple with a 
-cross-hair to show where to point the camera. The first picture is of the 
-actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, 
-figure \ref{bole}. Between these two pictures the camera will have to be tilted 
-vertically to capture the bole. The angle between these two camera positions is 
-recorded through the internal gyroscope. After this stage the user's role in 
-the application is done and the pictures is sent to one of SilviaTerra's web 
-servers. Here horizontal lines that run parallel with the bottom of the image 
-are drawn across the entire picture. These are drawn at different pixel heights 
-with equal pixel distance in between. The reason this is done is because 
-computer vision is still an experimental technology and is not fully capable 
-yet of being to detect the edges of trees with a background of forestry. 
-Seeing as this is a commercial product, it would be quite disadvantageous if 
-errors were commonplace. So instead the images with the newly drawn lines are 
-then forwarded to Amazon's Mechanical Turk service. Just like the name suggests 
-Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of 
-a automated chess board. The service is in its simplest form a crowd sourcing 
-internet marketplace where individuals or businesses can request tasks to be 
-completed. Usually the tasks are ones that are difficult for computers but 
-easy for humans. SilviaTerra creates tasks where people manually click where 
-the previously drawn lines, cross the edges of the tree. Ones of the points of 
-intersection are identified, their corresponding pixel coordinates are the sent 
-back to the web server. The pixel coordinates along with information from the 
-phone is piped into a program where the volume is then calculated. It is this 
-program that is the main topic of this research paper.
+After that they take two pictures.
+The interface is quite simple with a cross-hair to show where to point the camera.
+The first picture is of the actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{bole}.
+Between these two pictures the camera will have to be tilted vertically to capture the bole.
+The angle between these two camera positions is recorded through the internal gyroscope and accelerometer.
+After this stage the user's role in the application is done.
+Now the challenge is to extract useful information from these images.
+The pictures are sent to one of SilviaTerra's web servers.
+Here horizontal lines that run parallel with the bottom of the image are drawn across the entire picture.
+These are drawn at different pixel heights with equal pixel distance in between.
+The reason this is done is because computer vision is still an experimental technology and is computationally difficult as well as often being error-prone.
+SilviaTerra has instead opted a different route where the task of edge detection is performed manually by workers at Amazon's Mechanical Turk.
+So the images with the newly drawn lines are then forwarded to Mechanical Turk.
+The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed.
+Usually the tasks are ones that are difficult for computers but easy for humans.
+SilviaTerra creates tasks where people manually click where the previously drawn lines, cross the edges of the tree.
+Once the points of intersection are identified, their corresponding pixel coordinates are then paired with the phone sensor data and undergo photogrammetric analysis.
 
 \section{Method}
-Once all the required information has been retrieved it is piped into the 
-program that will be calculating the volume. The dataset that is used comprises 
-information about the phone and the results from Mechanical Turk. It is as follows:
+Once all the required information has been retrieved it is piped into the program that will be calculating the volume.
+The dataset that is used comprises information about the phone and the results from Mechanical Turk.
+It is as follows:
 \begin{itemize}
@@
 \subsection{Distance to tree}
-The DBH is used as a reference to find all the other measurements. The first
-unknown that must be discovered is the distance to the tree. This distance is 
-parallel with the ground. The difference between the x-coordinates of the left 
-and right pixel that correspond to the DBH is calculated, figure 
-\ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of
-the DBH is now known in both pixels and in a physical measurement, the ratio
-between the two can be determined by dividing the physical distance between 
-$X_1$ and $X_2$ with the number of pixels betwee them.
+The DBH is used as a reference to find all the other measurements.
+The first unknown that must be discovered is the distance to the tree.
+This distance is parallel with the ground.
+The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$.
+Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels between them.
+For example if the width of the tree is 100 pixels and its real life width is 50 centimeters then ratio is 0.5 cm per pixel.
 \begin{figure}[htp]
@@
 \end{figure}
- As seen in figure \ref{horizontal_triangle} an isosceles triangle can be 
- constructed with the horizontal angle of view as the vertex angle. The triangles 
- two congruent legs extend out towards the base of the triangle, $W_1$, which 
- is inline with the tree. To get the length of the base, $W_1$, the previously 
- calculated scale between pixels and the real world is multiplied with the 
- number of pixels in the horizontal resolution of the image. The distance to 
- the tree can easily calculated by drawing an extra line that bisects the vertex
- angle, contriving the angle $\theta$ which is exactly half of the angle of view 
- and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ 
- and futhermore producing a right angle triangle that can be solved. Finally to 
- actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
+As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constructed with the horizontal angle of view as the vertex angle.
+The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree.
+To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image.
+The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, resulting in an angle, $\theta$, which is exactly half of the angle of view and the actual distance to the tree $d$.
+This line also divides $W_1$ into $W_2$ and futhermore produces a right angle triangle that can be solved.
+Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
 
@@
 
-The next stage is retrieving the real world height to each set pixel coordinates 
-as well as the width between them, which will be the diameter of the tree. The 
-first step taken is to get the height. The vertical image plate scale which is 
-the ratio between degrees and pixels in the vertical axis is calculated by 
-dividing the vertical resolution of the image with the vertical angle of view. 
-This can be done because it is assumed that the silhouette of the tree is on a
-plane in the real world which corresponds to the image plane. With this method 
-it is possible to find out the angle to each pixel coordinate, $P$ in figure 
-\ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is 
-denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However 
-if the pixel coordinate was identified in the second image, when the phone was 
-tilted upwards, then angle between the distance to the tree and the lower boundary
-of the vertical angle of view, $B$, will need to be added, which is represented
-by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also 
-calculated so that the width can later be determined.
+The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree.
+The first step taken is to get the height.
+The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view.
+This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor.
+With this method it is possible to find out the angle to each pixel coordinate, example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
+This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}.
+However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}.
+The diagonal distance to $P$ is also calculated so that the width can later be determined.
 
-The final part of obataining the parameters for the set of pixel coordinates at 
-that height is to calculate the width between them. This is accomplished by 
-employing the method for calculating the distance to the tree, in reverse. 
-In this case what is known in figure \ref{horizontal_triangle} is the distance 
-$d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ 
-is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again 
-multiplied by 2 and then divided by the number of pixels in the horizontal 
-resolution of the image, once more returning the ratio between pixels and the 
-corresponding real world measurement. To finish off this ratio is the multiplied
-by the number of pixels between $X_1$ and $X_2$ dispensing the physical width
-of the tree at that certain height.
+The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them.
+This is accomplished by employing the method for calculating the distance to the tree, in reverse.
+In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
+$W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$.
+$W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
+To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.
 
 \subsection{Volume}
-Now the only thing left to produce the final result is to use these newly 
-calculated widths and heights to obtain the total volume. The current implemented 
-method of doing this is by integrating all the points using the trapezoidal rule.
-All the widths are divided by 2 to get the radius. Then they are squared and
-multiplied with $\pi$ to the obtain the basal area at each height. The basal 
-area is plotted along with their corresponding heights. The trapezoidal rule 
-then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
-The area between all the points is the summed together.
+Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume.
+The current implemented method of doing this is by integrating all the points using the trapezoidal rule.
+All the widths are divided by 2 to get the radius.
+Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height.
+The basal area is plotted along with their corresponding heights.
+The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
+The area between all the points is then summed together.
 
-This whole process is looped through for every identified pixel coordinate set 
-until the width and height is known for all of them. Once that is done, a new
-loop is created for calculating the total volume.
+This whole process is looped through for every identified pixel coordinate set until the width and height is known for all of them.
+Once that is done, a new loop is created for calculating the total volume.
 
 \subsection{Simulation}
-To test this method a \emph{virtual tree} was constructed. This tree was made 
-out of blue tape and tapered off so that the program could be thoroughly tested. 
-Along the tape points were marked out and the height to them plus the width
-between corresponding points were measured. Just like in the real application, 
-two pictures were taken. One of the DBH and one of the top of the tree. Each 
-set of points was then identified on the pictures and measured in pixels. The 
-parameters were then passed into the program and the results were cross-referenced
-with the real life measurements.
+To test this method a \emph{virtual tree}, that represented a 4m tall tree,was constructed.
+This tree was made out of blue tape and tapered off drastically so that the program could be thoroughly tested.
+Along the tape, points were marked out and the height to them plus the width between corresponding points was measured.
+Just like in the real application, two pictures were taken.
+One of the DBH and one of the top of the tree.
+Each set of points was then identified on the pictures and measured in pixels.
+The parameters were then passed into the program and the results were cross-referenced with the real life measurements.
 
@@
 		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
-		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401    & 445624    \\
+		452095   & 418053    & 428738    & 455700    & 451214    & 481471    & 501534    & 492573    \\
 		\hline
@@
 \subsection{Error tolerance}
-As seen in table \ref{heights} either when the vertical tilt angle is very wide, 
-standing close to the tree, or very narrow, standing far away, leads to error. I
-t's not clear what is causing this but there a two possibl reasons. The first one is
-that because of the short focal length of the camera and overall poor quality when 
-compared to a real camera, causes various forms of distrotion, especially around the 
-edges of the image. When the person is standing close to the tree so the vertical tilt 
-angle is over 45\degree, the points tend to lie exactly there, around the edges of 
-the image. The distortion then affects these points causing the pixel measurements 
-to be incorrect. The other problem occurs in the direct opposite situation, when 
-the person is standing far away and the angle is very narrow. Naturally what happens 
-is that the resolution of the camera just isn't high enough and the pixel measurements 
-just aren't accurate anymore. However most of this distortion seems to only affect the 
-vertical measurements because when looking at table \ref{widths} it is clearly seen that 
-the readings do not stray much at all from their actual values. Despite this the deviation 
-in the height readings still alters the volume calculations.
+As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error.
+It's not clear what is causing this but there a two possibl reasons.
+The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image.
+When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image.
+The distortion then affects these points causing the pixel measurements to be incorrect.
+The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow.
+Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore.
+However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values.
+Despite this the deviation in the height readings still alters the volume calculations.
+Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which is basically means they should stand somewhere between one tree length to two tree length's from the tree.
 \subsection{Limitations}
+As pointed out in the previous section the main hindrance attatched to this application is the actual camera. 
 \subsection{Future development}
-\subsection{Future development}
+The main area that could be expanded on is the edge detection.
+As explained in the background, today computer vision is still far from perfect.
+However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously.

### 2013-07-07 paper.tex
Added some more to discussion

@@
 As pointed out in the previous section the main hindrance attatched to this application is the actual camera.
+The problem is that most mobile phone cameras have a lot of different distortions that make it hard to make accurate measurements.
+Also when the tree is far away, the resolution of the camera can have a large affect on the final result.
+Human error is still large factor.
+If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect.
+Mechanical Turk is also a weak link because it is essentially humans doing the work and there nothing stopping them from also making mistakes.
 \subsection{Future development}
@@
 However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously.
+If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement.
+The possibilty might also be there in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.

### 2013-07-08 abstract.tex

% This is abstract.tex
Tree volume measurement is one of the most common ways to categorize trees in forest invetory. For a long time measuring trees has been done manually by us humans. Information collected is put through specialized stem-profile taper functions that need to be made for each species in each defined area. This method of collecting and analyzing data is labour intensive, costly and leaves huge margin for error. This research project aims to solve that and focuses on developing the mathematical basis and computational algorithms for a new mobile application project started by SilviaTerra. The application makes use of phone sensor data paired with photogrammetric analysis and has been successful in accurately calculating the volume of a tree when the person is standing at a reasonable distance from the tree.

### 2013-07-08 paper.tex

@@
 \section{Introduction}
-Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration.
+Tree volume measurement is one of the many parameters used when documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration.
 The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times.
@@
 Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
-Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches do not come without their downfalls.
+Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches does not come without it's downfalls.
 
@@
 
-SilviaTerra, a high-tech startup company, however is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results.
+SilviaTerra, a high-tech startup company,is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results.
 This is achieved through computer exactitude combined with statistical analysis.
@@
 
-The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree.
+The next stage is retrieving the real world height to each set of pixel coordinates as well as the width between them, which will be the diameter of the tree.
 The first step taken is to get the height.
@@
 This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor.
-With this method it is possible to find out the angle to each pixel coordinate, example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
+With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
 This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}.
@@
 
-The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them.
+The final part of obataining the parameters for each set of pixel coordinates is to calculate the width between them.
 This is accomplished by employing the method for calculating the distance to the tree, in reverse.
@@
 $W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
-To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.
+To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ resulting in the physical width of the tree at that certain height.
 
@@
 Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume.
-The current implemented method of doing this is by integrating all the points using the trapezoidal rule.
-All the widths are divided by 2 to get the radius.
-Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height.
-The basal area is plotted along with their corresponding heights.
-The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$.
-The area between all the points is then summed together.
+The new calculated heights and lengths are used to calculate a loess curve.
+A loess curve is a line that is fitted to the input points through local regression.
+The loess curve is then used to produce a polynomial spline function that can then be used to calculate the diameter at very small height intervals.
+At each new interval the volume can be calculated and is then added onto the current total volume.
+This will result in a much more accurate reading than by just numerically intergrating the original points.
 
@@
 As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error.
-It's not clear what is causing this but there a two possibl reasons.
+It's not clear what is causing this but there a two possible reasons.
 The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image.
@@
 Despite this the deviation in the height readings still alters the volume calculations.
-Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which is basically means they should stand somewhere between one tree length to two tree length's from the tree.
+Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which basically means they should stand somewhere between one tree length to two tree length's from the actual tree.
 \subsection{Limitations}
@@
 If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect.
-Mechanical Turk is also a weak link because it is essentially humans doing the work and there nothing stopping them from also making mistakes.
+Mechanical Turk is also a weak link because it is essentially humans doing the work and there is nothing stopping them from also making mistakes.
 \subsection{Future development}
@@
 If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement.
-The possibilty might also be there in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.
+The possibilty might also exist in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.

### 2013-07-08 paper.tex
Doing commits manually on Github now, git cli is just screwing around.

@@
 
-Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
+Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}.
 Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches does not come without it's downfalls.
@@
 However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}.
+To calculate the real world height to each set of pixel coordinates, the horizontal distance to the tree, $d$ in figure \ref{horizontal_triangle}, is multiplied wtih the angle to the pixel coordinate.
+In the DBH picture this angle is just $\theta$ and in the bole picture it is $\theta + \varphi$.
 The diagonal distance to $P$ is also calculated so that the width can later be determined.
@@
 However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values.
-Despite this the deviation in the height readings still alters the volume calculations.
+Despite this the deviation in the height readings still alters the volume calculations, as observed in table \ref{volumes}.
 Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which basically means they should stand somewhere between one tree length to two tree length's from the actual tree.
@@
 Mechanical Turk is also a weak link because it is essentially humans doing the work and there is nothing stopping them from also making mistakes.
+Although the application has removed a lot of possible human error it has also introduced new machine error.
+Sensor data from the phone can be inaccurate.
+The built-in gyroscope and accelerometer can sometimes give very rough measurements and a few degrees can change the whole outcome.
+Plus if they are not callibrate properly then that will also lead to error.
 \subsection{Future development}
@@
 If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement.
-The possibilty might also exist in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.
+The possibilty might also exist in the future to have the application calculate extra things like canopy area and volume of external branches and limbs.

### 2013-07-08 paper.tex
Minor edits

@@
 These taper functions are based on numerous measurements and have been refined over the years.
-The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions and are usually custom made for a certain species in a certain area.
+The main problem with taper functions is that to get an accurate result they need to be callibrated to hundreds of local conditions and are usually custom made for a certain species in a certain area.
 This again is cost prohibitive and only so much of a forest can be accounted for in the taper function.
@@
 The data can then later be used with SilviaTerra's other products.
-If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, \ref{input}.
+If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, figure \ref{input}.
 \begin{figure}[!htb]
@@
 So the images with the newly drawn lines are then forwarded to Mechanical Turk.
-The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed.
+The service in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed.
 Usually the tasks are ones that are difficult for computers but easy for humans.
@@
 The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$.
-Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels between them.
+Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement (from figure \ref{input} where the user typed it in), the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels between them.
 For example if the width of the tree is 100 pixels and its real life width is 50 centimeters then ratio is 0.5 cm per pixel.
@@
 The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view.
+For example if the vertical angle of view is 60 degrees and there are 3000 pixels in the vertical resolution then the ratio between the two is 50 pixels per degree.
 This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor.
+Basically the tree doesn't bend towards or from the camera.
 With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
@@
 This is accomplished by employing the method for calculating the distance to the tree, in reverse.
-In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
+In this case what is known in figure \ref{horizontal_triangle} is the distance $d$, which was the diagonal distance to P, and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
 $W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$.
-$W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
+$W_2$ is then again multiplied by 2, to give $W_1$, and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
 To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ resulting in the physical width of the tree at that certain height.

### 2013-07-08 paper.tex
Added equations to method section and reworked the text.

@@
 Tree volume measurement is one of the many parameters used when documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration.
-The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times.
+The need for information about the natural resources of the Earth, including tree volume, has been increasing exponentially in recent times.
 To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.
@@
 \subsection{Distance to tree}
+\subsubsection{Explanation}
 The DBH is used as a reference to find all the other measurements.
@@
 The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree.
-To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image.
+To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image, Eq.
+\ref{w1}.
 The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, resulting in an angle, $\theta$, which is exactly half of the angle of view and the actual distance to the tree $d$.
-This line also divides $W_1$ into $W_2$ and futhermore produces a right angle triangle that can be solved.
-Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.
+This line also divides $W_1$ into $W_2$, Eq.
+\ref{w2}, and futhermore produces a right angle triangle that can be solved.
+Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$, Eq.
+\ref{d}.
+
+ \subsubsection{Equations}
+ \begin{itemize}
+	\item $Hr$: Horizontal resolution.
+	\item $DBH$: Diameter at Breast Height.
+\end{itemize}
+
+\begin{equation}
+	W_1 = \frac{DBH}{(X_1 - X_2)} \times Hr
+	\label{w1}
+\end{equation}
+\begin{equation}
+	W_2 = \frac{W_1}{2}
+	\label{w2}
+\end{equation}
+\begin{equation}
+	d = \frac{W_2}{\tan{\theta}}
+	\label{d}
+\end{equation}
 
 \subsection{Heights and widths} 
+\subsubsection{Explanation}
 \begin{figure}[!htb]
@@
 The first step taken is to get the height.
-The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view.
+The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view, Eq.
+\ref{ips}.
 For example if the vertical angle of view is 60 degrees and there are 3000 pixels in the vertical resolution then the ratio between the two is 50 pixels per degree.
@@
 Basically the tree doesn't bend towards or from the camera.
-With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view.
+With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view by dividing its Y position by the image plate scale, Eq.
+\ref{theta}.
 This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}.
-However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}.
-To calculate the real world height to each set of pixel coordinates, the horizontal distance to the tree, $d$ in figure \ref{horizontal_triangle}, is multiplied wtih the angle to the pixel coordinate.
-In the DBH picture this angle is just $\theta$ and in the bole picture it is $\theta + \varphi$.
-The diagonal distance to $P$ is also calculated so that the width can later be determined.
+However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}, Eq.
+\ref{varphi}.
+To calculate the real world height to each set of pixel coordinates, the horizontal distance to the tree, $d$ in figure \ref{horizontal_triangle}, is multiplied with tan of the angle.
+In the DBH picture this angle is just $\theta$ and in the bole picture it is $\theta + \varphi$, Eq.
+\ref{py1} or Eq.
+\ref{py2}.
+The diagonal distance to $P$ is also calculated so that the width can later be determined, Eq.
+\ref{Dd1} or Eq.
+\ref{Dd2}.
 
@@
 This is accomplished by employing the method for calculating the distance to the tree, in reverse.
-In this case what is known in figure \ref{horizontal_triangle} is the distance $d$, which was the diagonal distance to P, and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
-$W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$.
+In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ which is actually the diagonal distance to $P$, and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$.
+$W_2$ is calculated by multiplying $d$ with $\tan{\theta}$, Eq.
+\ref{w2a}.
 $W_2$ is then again multiplied by 2, to give $W_1$, and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement.
-To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ resulting in the physical width of the tree at that certain height.
+To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ resulting in the physical width of the tree at that certain height, Eq.
+\ref{widtheq}.
+
+\subsubsection{Equations}
+\begin{itemize}
+	\item $Vav$: Vertical angle of view
+	\item $Vr$: Vertical resolution
+	\item $Ips$: Image plate scale
+	\item $H$: Height to pixel coordinate
+	\item $Vta$: Vertical tilt angle (pitch)
+	\item $Dd$: Diagonal distance to $P$
+\end{itemize}
+
+\begin{equation}
+	Ips = \frac{Vr}{Vav}
+	\label{ips}
+\end{equation}
+\begin{equation}
+	\theta = \frac{Py}{Ips}
+	\label{theta}
+\end{equation}
+\begin{equation}
+	\varphi = Vta - Vav
+	\label{varphi}
+\end{equation}
+\begin{equation}
+	H = d \times \tan{\theta}
+	\label{py1}
+\end{equation}
+\begin{equation}
+	H = d \times \tan{(\theta + \varphi)}
+	\label{py2}
+\end{equation}
+\begin{equation}
+	Dd = \frac{d}{\cos{\theta}}
+	\label{Dd1}
+\end{equation}
+\begin{equation}
+	Dd = \frac{d}{\cos{(\theta + \varphi)}}
+	\label{Dd2}
+\end{equation}
+\begin{equation}
+	W_2 = d \times \tan{\theta}
+	\label{w2a}
+\end{equation}
+\begin{equation}
+	\text{Width of tree} = \frac{W_2 \times 2}{Hr} \times (X_2 - X_1)
+	\label{widtheq}
+\end{equation}
 
